R Notebook

### 2017-07-03 doc/RMD/wmn-differentiation.Rmd
add amova table and density plot

```{r correction}
dat11
dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
dat11cc
```

@@

```{r AMOVA-Region, cache = TRUE}
(ssc_amova_region <- poppr.amova(dat11cc, ~SourceType/Region, dist = bd, quiet = TRUE))
ssc_amova_region_test <- randtest(ssc_amova_region, nrepet = 999)
plot(ssc_amova_region_test)
ssc_amova_region_test
```

@@

```{r AMOVA-nomex, cache = TRUE}
datnomex <- setPop(dat11cc, ~Region) %>% popsub(blacklist = "Mexico")
bdnm     <- bruvo.dist(datnomex, replen = other(datnomex)$REPLEN)
(ssc_amova_nm <- poppr.amova(datnomex, ~SourceType/Region, dist = bdnm, quiet = TRUE))
ssc_amova_nm_test <- randtest(ssc_amova_nm, nrepet = 999)
plot(ssc_amova_nm_test)
ssc_amova_nm_test
```

@@

>>>>

```{r correction}
dat11
dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
dat11cc

make_amova_table <- function(am, amt, samples = "Region"){
  tot <- nrow(am$results)
  res <- data.frame(list(am$results[-tot, c("Df", "Sum Sq")], 
                         Percent = am$componentsofcovariance[-tot, 2],
                         Pval    = rev(amt$pvalue), 
                         Phi     = am$statphi$Phi[-tot]))
  res <- as.matrix(res)
  colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
                     "Phi statistic")
  names(dimnames(res)) <- c("levels", "statistic")
  rownames(res) <- gsub("samples", samples, rownames(res))
  return(res)
}

make_amova_printable <- function(amtab, amtabcc){
  am_array <- array(dim      = c(dim(amtab), 2),
                    dimnames = c(dimnames(amtab), 
                                 list(c("full", "clone-corrected"))))
  am_array[, , 1] <- amtab
  am_array[, , 2] <- amtabcc
  tabfun <- function(x){
    x <- paste0(paste0(signif(x, 3), collapse = " ("), ")")
    return(x)
  }
  res <- apply(am_array, c(1, 2), tabfun)
  return(res)
}
```

@@

```{r AMOVA-Region, cache = TRUE}
ssc_amova_region <- poppr.amova(dat11cc, ~SourceType/Region, dist = bd, quiet = TRUE)
ssc_amova_region_test <- randtest(ssc_amova_region, nrepet = 999)
plot(ssc_amova_region_test)
(ssc_amova_table <- make_amova_table(ssc_amova_region, ssc_amova_region_test))
```

@@

```{r AMOVA-nomex, cache = TRUE}
datnomex <- setPop(dat11cc, ~Region) %>% popsub(blacklist = "Mexico")
bdnm     <- bruvo.dist(datnomex, replen = other(datnomex)$REPLEN)
ssc_amova_nm <- poppr.amova(datnomex, ~SourceType/Region, dist = bdnm, quiet = TRUE)
ssc_amova_nm_test <- randtest(ssc_amova_nm, nrepet = 999)
plot(ssc_amova_nm_test)
(ssc_amova_nm_table <- make_amova_table(ssc_amova_nm, ssc_amova_nm_test))
```

```{r AMOVA_table, comment = NA}
make_amova_printable(ssc_amova_table, ssc_amova_nm_table) %>%  
  as_tibble() %>%
  add_column(Hierarchy = c("Between Source", "Between Region within Source", "Within Region"), .before = 1) %>%
  rename(ps = `Phi statistic`) %>%
  mutate(ps = gsub("0\\.00(\\d{1})(\\d{2})", "\\1.\\2e^-3^", ps)) %>%
  rename(`$\\Phi statistic$` = ps) %>%
  rename(`% variation` = `Percent variation`) %>%
  rename(S.S. = `Sum of Squares`) %>%
  select(-P) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>% 
  huxtable::set_col_width(c(1.3, 0.7, 0.9, 1, 1.3)) %>% 
  huxtable::set_align(huxtable::everywhere, 2:5, "center") %>% 
  huxtable::print_md(max_width = 95)
```

@@

## Visualizing the partitions

We can visualize the partitions if we create distributions showing the genetic
distance.

```{r distrib, fig.width = 6, fig.height = 3}
wmn_inds <- (setPop(dat11cc, ~SourceType) %>% pop()) == "wmn"

# Function to set the upper triangle to NA to avoid over-representation.
set_upper_tri_NA <- function(d){d[upper.tri(d)] <- NA; diag(d) <- NA; d}

# dist data converted to matrix fed into here. Removes NAs from previous function
tidy_dist <- . %>%
  as.data.frame() %>% 
  rownames_to_column("from") %>% 
  gather(to, distance, -from) %>% 
  filter(!is.na(distance))
wmn_distance   <- as.matrix(bd)[wmn_inds, wmn_inds] %>% set_upper_tri_NA() %>% tidy_dist
nwmn_distance  <- as.matrix(bd)[!wmn_inds, !wmn_inds] %>% set_upper_tri_NA() %>% tidy_dist
inter_distance <- as.matrix(bd)[wmn_inds, !wmn_inds] %>% tidy_dist

dists <- bind_rows(`White Mold Nurseries` = wmn_distance, 
                   `Between` = inter_distance,
                   `Other Sources` = nwmn_distance, 
                   .id = "Comparison") %>%
  mutate(Comparison = forcats::fct_inorder(Comparison))
dists
ggplot(dists, aes(y = distance, x = Comparison, fill = Comparison)) +
  geom_violin() +
  geom_boxplot(width = 0.25) +
  scale_fill_manual(values =  c(`Other Sources` = "grey35", `White Mold Nurseries` = "grey95", Between = "grey65")) +
  scale_color_manual(values = c(`Other Sources` = "grey35", `White Mold Nurseries` = "grey95", Between = "grey65")) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/2) +
  theme(legend.position = "none") +
  theme(axis.title.x = element_blank()) +
  theme(axis.text = element_text(color = "black")) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50")) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank()) +
  ylab("Bruvo's Distance")
```

```{r dist_comparison, results = "asis"}
summary_dists <- dists %>% 
  group_by(Comparison) %>% 
  summarize(`Average Distance` = mean(distance),
            `Standard Deviation` = sd(distance)) 
knitr::kable(summary_dists, digits = 3)
```

On first look, it appears that there's not much difference. Looking at the
table, we can see that the average distance within White Mold Nurseries is 
`r round(summary_dists[["Average Distance"]][1], 3)`, whereas the average distance within Other
Sources is `r round(summary_dists[["Average Distance"]][3], 3)`. This is well within the standard
deviation for both distributions, but we also recognize that a single step of
Bruvo's distance for 11 loci in a haploid organism is `r round(0.5/11, 3)`.

# Summary
